\documentclass[12pt]{article}
\usepackage{fullpage}
%\usepackage{mathptmx}
\usepackage{times}
\usepackage{ifthen}
\usepackage{microtype}
\usepackage{color}
\usepackage{xspace}
\usepackage{makeidx}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage[normalem]{ulem}
\usepackage{titlesec}
\usepackage{tikz}
\usepackage[plainpages=false,debug]{hyperref} % should be last package

%\usepackage{fancyhdr}
%
%\fancyfoot{}
%\fancyhead[RO,LE]{\thepage}
%\fancyhead[LO]{}
%\fancyhead[RE]{}

\setlist{nosep} % tighten itemized lists

\definecolor{diColor}{rgb}{0,0,0.8}
\definecolor{deleteColor}{rgb}{1,0,0}
\definecolor{addColor}{rgb}{0.0,0.5,0}

\definecolor{fixmeColor}{rgb}{0.8,0,0} % hide them as black

\newcommand\delete[1]{\color{deleteColor}\sout{\textbf{#1}}\color{black}\xspace%
\marginpar{\color{deleteColor}$\Delta$\color{black}}}

\newcommand\add[1]{\color{addColor}\textbf{#1}\color{black}\xspace%
\marginpar{\color{addColor}$\Delta$\color{black}}}

%\newcommand\edited{\marginpar{\color{red}$\Delta$\color{black}}}

\newcommand{\course}[2][EMPTY]%
    {\ifthenelse%
        {\equal{#1}{EMPTY}}%
        {OCEA #2}%
        {#1 #2}%
    }

\newcommand{\di}[2][EMPTY]%
    {\ifthenelse%
        {\equal{#1}{EMPTY}}%
        {\color{diColor}#2\color{black}\index{#2}}%
        {\color{diColor}#2\color{black}\index{#1}}%
    } 

\newcommand{\fixme}[1]{\color{fixmeColor}$\{$#1$\}$\color{black}\index{$>>>>$FIXME$<<<<$}}

%\newcommand{\npara}[1]{\noindent\textbf{\thesubsection.#1}} % for numbered paragraphs of appeals section

\setlength{\parindent}{0em}
% number paragraphs
\newcommand{\parnum}{\arabic{parcount}}
\newcounter{parcount}
%\newcommand\p{\stepcounter{parcount}\leavevmode[\parnum]\hspace{0.2em}\marginpar[\hfill\parnum]{\parnum}}
\newcommand\p{\stepcounter{parcount}\leavevmode{\small[\parnum]}\hspace{0.2em}}
\newcommand\cp{\setcounter{parcount}{0}}


% Use macros to get indexing, and ensure uniform representation
\newcommand{\curcom}{\index{Committees!Curriculum}Curriculum Committee\xspace}
\newcommand{\advcom}{\index{Committees!Advisory}Advisory Committee\xspace}
\newcommand{\gocom}{\index{Committees!Graduate Oversight}Graduate Oversight Committee\xspace}
%\newcommand{\GC}{\index{Graduate Coordinator}Graduate Coordinator\xspace}
\newcommand{\GC}{Graduate Coordinator\xspace}
%\newcommand{\GS}{\index{Graduate Secretary}Graduate Secretary\xspace}
\newcommand{\GS}{Graduate Secretary\xspace}
\newcommand{\formAPR}{\index{forms!Annual Progress Report}Annual Progress Report Form\xspace}
\newcommand{\formGSP}{\index{forms!Graduate Student Program}Graduate Student Program Form\xspace}
\newcommand{\formGSPU}{\index{forms!Graduate Student Program}Graduate Student Program Update Form\xspace}
\newcommand{\formSPRF}{\index{forms!Supplementary Program Report}Supplementary Program Report Form\xspace}
\newcommand{\formST}{\index{forms!Sea Time}Sea Time Form\xspace}
\newcommand{\FGS}{\index{Faculty of Graduate Studies (FGS)}FGS\xspace}

\setlength{\parskip}{0.7ex plus 0.5ex minus 0.2ex}

\setcounter{secnumdepth}{5}
\titleformat*{\section}{\large\bfseries}
\titlespacing*{\section}{0pt}{*0}{0pt}
\titleformat*{\subsection}{\bfseries\slshape}
\titlespacing*{\subsection}{0pt}{1.5ex plus 0.5ex minus 0.2ex}{0pt}
\titleformat*{\subsubsection}{\slshape}
\titlespacing*{\subsubsection}{0pt}{*0}{0pt}
%\titleformat*{\subsubsubsection}{}
%\titlespacing*{\subsubsubsection}{0pt}{*0}{0pt}
\titleformat*{\paragraph}{\slshape}
\titlespacing*{\paragraph}{0pt}{*0}{0pt}

\newcommand{\ExternalLink}{%
    \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{% 
        \begin{scope}[x=1ex, y=1ex]
            \clip (-0.1,-0.1) 
                --++ (-0, 1.2) 
                --++ (0.6, 0) 
                --++ (0, -0.6) 
                --++ (0.6, 0) 
                --++ (0, -1);
            \path[draw, 
                line width = 0.5, 
                rounded corners=0.5] 
                (0,0) rectangle (1,1);
        \end{scope}
        \path[draw, line width = 0.5] (0.5, 0.5) 
            -- (1, 1);
        \path[draw, line width = 0.5] (0.6, 1) 
            -- (1, 1) -- (1, 0.6);
        }
    }

%\renewcommand*\contentsname{}

\makeindex

\title{Oceanography Graduate Handbook}
\author{}
\date{September, 2016}


\begin{document}
\clearpage
\maketitle
\thispagestyle{empty}

\textbf{Disclaimer.} The guidelines in this handbook supplement, but do not,
and cannot, reduce the requirements listed in the rules and regulations
published in the Graduate Studies Calendar of Dalhousie University.  As such,
this document adds further requirements that students must accomplish in order
to complete a degree program in Oceanography.  Such additional prerequisites
are permitted by Graduate Studies.  In the event of a real or perceived
conflict in such stipulations, Graduate Studies (\FGS) requirements and rules
shall prevail.

\newpage

\renewcommand{\baselinestretch}{0.5}\normalsize

\pagenumbering{roman}
{\footnotesize
\setcounter{tocdepth}{2}
\tableofcontents
}
\thispagestyle{empty}
\renewcommand{\baselinestretch}{1.0}\normalsize

\pagenumbering{arabic}
%\pagestyle{fancy}
\setcounter{page}{0}

\section{\label{sec:glossary}Glossary of acronyms and terms}

The following list explains some acronyms and special terms used in this
document.

\begin{itemize}

    \item ``CRN'' refers to a Course Registration Number. For example, the
        Physical Oceanography core class, listed as \course{5120} in the University
        Calendar, has CRN 12070. Consult the Registrar's website\footnote{Registrars website of Oceanography classes \href{https://dalonline.dal.ca/PROD/fysktime.P_DisplaySchedule?s_term=201710,201720&s_subj=OCEA&s_district=100}{\ExternalLink}.}

    \item ``External'' \fixme{write a few lines here, with the 3 types
        discussed in the meeting of 20160527}.

    \item ``\FGS''  stands for Faculty of Graduate Studies.

    \item ``NSERC'' stands for Natural Science and Engineering Research
        Council, a funding agency.

\end{itemize}


% A.	GENERAL  PROGRAM REQUIREMENTS
\section{\label{sec:general_program_requirements}General Program Requirements}

%A1	REGISTRATION
\subsection{Registration}

\p Graduate students must maintain their registration in all three terms of
each year in their program, except in cases where a formal \di[leave of
absence]{Leave of Absence} has been approved by the Faculty of Graduate
Studies. This involves registering for a ``course'' named ``Registration
Course--Graduate'' (designated \course[REGN]{9999}, and also called a
``fee-generating course''), as well as either ``Masters Thesis''
(\course{9000}) or ``PhD thesis'' (\course{9530}); see the
Registrar's website\footnote{FGS Registration Information website
\href{http://www.dal.ca/faculty/gradstudies/currentstudents/registration.html}{\ExternalLink}.}
for more information, including the relevant Course Registration Numbers. It is
critical to register for these courses in time, since failure to register at
lease 1 month prior to the beginning of a term will result in non-payment of
scholarships and stipends.


%\p Registration consists of
%\begin{enumerate}
%    \item REGN 9999
%    \item Thesis Code
%    \item Courses (if applicable)
%\end{enumerate}

\p If \course{REGN~9999} is not added for each term, graduate students are not
considered to be registered. Failure to register at least one month prior to
the beginning of each term will result in non-payment of scholarships and
stipends.

\p Students who fail to register within the approved deadlines will be
considered to have \di{lapsed registration}. Such students will not be
permitted to submit a thesis, nor will they receive any services from the
University during that academic term. Students who allow their registration to
lapse will be considered to have withdrawn and will be required to apply for
readmission\footnote{FGS Regulations, section 5.2.3: Failure to Register
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10848&loaduseredits=False}{\ExternalLink}.}.

%%A1.1	REGN 9999
%\subsection{REGN 9999}
%
%\cp
%
%\p Students must register at least one month prior to the beginning of each term.
%
%\p Students must register for the Fee Generating
%Course\index{courses!fee-generating} (REGN 9999) in all three terms for the
%duration of their program. If REGN 9999 is not added for each term, graduate
%students are not considered to be registered. Failure to register at least one
%month prior to the beginning of each term will result in non- payment of
%scholarships and stipends.

%%A1.2	THESIS CODE
%\subsection{Thesis Code}
%
%\cp
%
%\p Students must register at least one month prior to the beginning of each term.
%
%\p Students must register for the Thesis Code in all three terms for the duration
%of their program. Failure to do so in a term during which formal classes are
%not being taken will result in a blank term on the student's transcript, which
%means that there will be no documentation demonstrating that work was done on
%the thesis.
%
%\p The CRN for each term can be found in the Academic Timetable under the
%Oceanography course listings (MSc – OCEA 9000, PhD – OCEA 9530).



% A2	COURSES
\subsection{\label{sec:courses}Courses}

\cp

\p In addition to the artificial courses listed in the previous section,
students take conventional courses, chosen in consultation with their
supervisors, and pursuant to the requirements set out in
section~\ref{sec:msc_course_requirements} or
section~\ref{sec:phd_course_requirements} for MSc and PhD students,
respectively.

\p Within the Department of Oceanography, courses are designated as either
``core'' or ``non-core.'' The core classes are
Biological Oceanography (\course{5140}),
Chemical Oceanography (\course{5130}),
Geological Oceanography (\course{5110}),
and
Physical Oceanography (\course{5120}). These 
are broadly based courses that are intended to
help students gain a wide understanding of the major areas of oceanography.
They are offered every year, typically with Physical and Chemical in the 
September term, to set the stage for Geological and Biological in the January term.
Note that these courses are cross-listed to 4th year classes at the 
Undergraduate level, and there are circumstances in which they may be waived
for Dalhousie students who have taken them at this lower level.\fixme{cite section numbers}

\p The non-core classes tend to be more advanced and specialized. They include
Fluid Dynamics (\course{5311}),
Time Series Analysis (\course{5210}),
Ocean Dynamics (\course{5221}),
Estuary, Coast and Shelf Dynamics (\course{5222}),
Numerical Modelling (\course{5220}),
Ocean Waves (\course{5223}),
Introduction to Acoustical Oceanography (\course{5250}),
Advanced Chemical Oceanography (\course{5290}),
and
Advanced Marine Particles (\course{5293}).  Not all of these are offered
every year. Details of offerings are available in the departmental office 
and at the Registrar's website\footnote{Academic Timetable for Oceanography
\href{https://dalonline.dal.ca/PROD/fysktime.P_DisplaySchedule?s_term=201710,201720&s_subj=OCEA&s_district=100}{\ExternalLink}.}.


\p Students select courses in consultation with their supervisors, with the
approval of members of their \advcom. Students must register for the 5000 level
courses (4000 level is for undergraduates only) prior to the registration
deadline\footnote{Course registration deadlines can be found in the Graduate
Studies Academic Calendar
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2677&loaduseredits=False}{\ExternalLink}.}.


\p A \formGSP must be submitted to the \GC within one month of the start of
your program.  Students must submit a \formGSPU to the \GS when a decision to
make changes to course requirements is made\footnote{Graduate Student Program
Forms can be downloaded from the FGS website
\href{http://www.dal.ca/faculty/gradstudies/currentstudents/forms.html}{\ExternalLink}.}.
Failure to do so can result in the assessment of additional fees for courses
not included in your program requirements, or affect the student's ability to
graduate.

%\p In addition to the core classes as described above, there are several
%classes designed to provide students with the tools and background knowledge
%appropriate to their particular area of research. As explained in sections B
%and C, the detailed requirements differ for MSc and PhD and between the
%sub-disciplines of Oceanography. The choice of classes is made by the student
%in consultation with the thesis supervisor, with the approval of the advisory
%committee. Students must register for the 5000 level courses (4000 level is for
%undergraduates only) prior to the registration deadline\footnote{Course
%registration deadlines can be found in the Graduate Studies Academic Calendar
%\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2677&loaduseredits=False}{\ExternalLink}.}.

%A2.3	GRADING
%\subsubsection{Grading}

\p FGS Regulations\footnote{FGS Regulations, section 7.6: Classes and Grades
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10850&loaduseredits=False}{\ExternalLink}.}
stipulate that graduate students must achieve a grade of B- or higher in all
classes required as part of their degree program. The only grade that can be assigned
below a B- is an F. Any student who receives an F
will be withdrawn
from the Graduate Program and must apply for re-admission. Such a student may
apply, in writing, to the department for immediate reinstatement\footnote{FGS
Regulations, section~5.4.1: Readmission of Students
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10848&loaduseredits=False}{\ExternalLink}.}.
Reinstatement to a program after a failing grade must be supported by the
\GC, and must be approved in writing by the Faculty of
Graduate Studies. If readmitted, any subsequent 'F' will result in a final
program dismissal. Any academic withdrawal and reinstatement will be recorded
on the student's official transcript.
% above was 5.2.6 but now seems to be 5.4

%A2.4	REGISTERING FOR COURSES AT ANOTHER INSTITUTION
%\subsection{Registering for courses at another institution}

%\cp

\p Classes approved by the Department and Faculty of Graduate Studies (after
examination of class descriptions\fixme{Q: who approves these? I've done that
as the professor for my own course. Is that the rule, or the habit?}) can be
taken at other universities as part of the graduate degree program, provided
the class is not available at Dalhousie\footnote{FGS Regulations, section
7.6.6: Letters of Permission
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10850&loaduseredits=False}{\ExternalLink}.}.
The Letter of Permission Form\footnote{FGS forms and documents: Letter of
Permission--Request Form
\href{http://www.dal.ca/faculty/gradstudies/currentstudents/forms.html}{\ExternalLink}.}
must be completed and submitted to the \GC in advance. Such approval will not
be given retroactively.


%A3	THESIS PROPOSAL
\subsection{Thesis proposal}

\cp

%\p When a particular research area has been identified and an advisory committee
%appointed, students are required to provide their advisory committees with
%detailed thesis proposals. This document, typically developed in collaboration
%with the supervisor, should demonstrate the student's
%
%\begin{itemize}
%\item Background of appropriate literature
%\item Awareness of current research activity
%\item Ability to formulate pertinent scientific hypotheses
%\item Appreciation of the time, effort, and resources necessary to achieve the thesis objectives
%\end{itemize}

%\p The acceptance of a thesis proposal is a critical step in a student's program.

\p See sections~\ref{sec:msc_thesis_proposal} and~\ref{sec:phd_thesis_proposal}
for requirements in the MSc and PhD programs, respectively.

%A4	SEA TIME
\subsection{Sea time}

\cp

% \p Graduate students in Oceanography are required to spend time at sea on
% oceanographic ships to familiarize themselves with oceanographic techniques,
% even if their research does not require measurements at sea. Students are
% required to submit a \di[sea time]{Sea Time} Form to the Graduate Secretary.
% The \curcom must approve completed forms. Requests for special consideration by
% the \curcom for \di{fieldwork} in lieu of sea time may be made in extenuating
% circumstances.

\p The Department of Oceanography has a longstanding tradition of requiring our
graduate students to go to sea\index{sea time} on research cruises as part of
their degree requirements. This tradition meets several objectives of our
graduate education:

\begin{itemize}

    \item Sea time directly exposes students of all backgrounds and research
        interests to the rewards and challenges of gathering data at sea;

    \item Sea time offers the opportunity to network with other scientists in a
        unique and stimulating environment.

    \item Sea time provides practical education about how to solve problems that
        arise when outside assistance is not readily at hand;

    \item \di[sea time]{Sea time} emphasizes the importance of practicing
        expeditionary behaviour whereby adherence to safe procedures and
        maintenance of clear and open communication are paramount;

\end{itemize}

\p Arrangements to meet the \di{sea time} requirement are the responsibility of
the student in consultation with the supervisor. Financial costs associated
with meeting the \di{sea time} requirement should be borne either by the
supervisor, a travel grant, or an outside investigator funded to participate in
the research cruise. In general \fixme{is this phrase ok? Another choice might
be ``normally'' or ``usually.'' I think CTT has expressed opinions on how to
phrase such things.} ``sea time'' means spending 3 or more nights at sea on a
research vessel that is conducting active research. Students are required to
submit a \formST to the \GS after completion of this requirement, and the
\curcom must approve completed forms.

\p In a limited set of extenuating circumstances, students may have the
sea-time requirement waived and replaced with commensurate experience. For
example, students facing physical or mental challenges to working at sea,
students with extensive experience at sea prior to coming to Dalhousie, or
students with demanding nearshore/intertidal field research projects that limit
time available to go to sea may be granted a waiver of the sea-time
requirement. To apply for a waiver, the student composes a letter to the
\curcom stating the reason for the request and proposing commensurate
replacement experience. An accompanying letter supporting the student’s request
must be composed by the supervisor and signed by the supervisory committee. The
decision to grant or deny the request for a sea time waiver is the
responsibility of the \curcom.


% A5	SEMINARS
\subsection{Seminars}

\p Students are required to attend the general \di{departmental
seminars}\index{seminars} and to attend, and participate in, the specialty
seminars in their field of interest.  Students who are unable to attend
seminars regularly must have the specific agreement of their \advcom that this
requirement is waived and this must be communicated to the departmental office
in a written memo signed by the supervisor. It is important to note that
materials presented in these seminars may form part of the questions for the
PhD Qualifying Exam and the PhD Thesis Proposal Defence.



 

%A6	CHANGE IN STATUS
\subsection{Status}

\cp

\p Any change in status, such as transfer from MSc to PhD (see
section~\ref{sec:transfer_to_phd}), a leave of absence, and entrance to another
degree program must be recommended by the \advcom to the \gocom.  Final
approval for a change in status will be made by the \gocom, which will meet as
required to review these requests.  Students should consult the Graduate
Secretary regarding the required documentation.


%A7	ANNUAL PROGRESS REPORT
\subsection{Annual progress report}

\cp

\p Students must submit an \formAPR each year, one month prior to the
anniversary of their admission date. Failure to submit this report will result
in delays in registration and funding.

\p This report is made with the Graduate Student Information System (GSIS)
through DalOnline\footnote{DalOnline
\href{https://dalonline.dal.ca}{\ExternalLink}.}.  The supervisor and the \GC
must approve this report electronically, and students should bear this in mind
so that the deadline can be met.

\p A hard copy of the Oceanography \formSPRF (available from the GS) must also
be approved by your supervisor and the \GC, and submitted to the \GS annually,
following the same deadline as the \formAPR.

\p The \formSPRF is intended to be saved electronically and added to on a
regular basis, similar to a CV. An up-to-date copy of the \formSPRF is to be
brought to all meetings with the \GC.


%A8	TIME LIMITS FOR COMPLETION OF DEGREES
\subsection{Time limits for completion of degrees}

\cp

\p Graduate students have a maximum period of time within which to complete their
graduate program, 4 years for MSc and 6 years for PhD.  Extensions may be
granted by FGS on the recommendation of the department, along with a
satisfactory Progress Report\footnote{FGS Regulations, section 7.1: Maximum
Time for Degree Completion and Extensions
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10850&loaduseredits=False}{\ExternalLink}.}
Under no circumstances can a student be registered in a program beyond 10 years
from their initial registration in the program.

\section{Supervision}

%A9.3	SUPERVISOR
\subsection{Supervisor}

\cp

\p The Supervisor provides direct supervision of the student's research and is expected to
\begin{itemize}
\item Advise the student on the choice of a research topic, on the possible directions to emphasize during the work, and the point at which it should be concluded.
\item Supervise the research, as well as the preparation of the proposal, progress reports and thesis.
\item Ensure that the resources necessary to the thesis project are made available and that any necessary skills are acquired by the student.
\item Assess the student's progress, evaluating strengths and weaknesses.
\item Provide current updates of progress to the student's file.
\item Ensure that the student is funded
\item Select thesis examining committee members (following the guidelines for composition of the various examining committees).
\end{itemize}

\p A supervisor is changed only with the approval of the Department. This may
arise by mutual agreement within the advisory committee or at the request of
the student (to either the \GC or the Chair of the Department).

\p In order to take advantage of the special expertise and facilities available,
the role of supervisor is not restricted to permanent faculty members of the
Department. However, supervisors must be Adjunct Professors or Full Professors
with Faculty of Graduate Studies membership in Oceanography.\fixme{check on the
rules, probably citing an FGS table}

\subsection{Internal supervisor}
\cp

\p In the case where a student has a supervisor who is not a full time faculty
member in the Department of Oceanography, there must be an internal supervisor
who is a faculty member in the Department. The responsibilities of the
\di{internal supervisor} include:

\begin{itemize}

\item Advising on academic requirements, course load, waiver of courses,
    general departmental affairs, etc.

\item Ensuring that the student has suitable office and laboratory space.

\item In consultation with the student, selecting a supervisor and advisory
    committee appropriate to the student's research interests (In many cases
    the chair will also be the supervisor.)\fixme{who is the chair?}

\item Ensuring that adequate reports and records of committee meetings are kept
    and submitted to the appropriate departmental files, and that necessary
    decisions as to the academic program, (e.g. thesis proposal), are made on
    an appropriate time-scale.

\item Assessing the progress of the student both in academic work and research.

\end {itemize}
 

%B.	MSc  PROGRAM REQUIREMENTS




%A9	THESIS ADVISORY COMMITTEE
\subsection{Advisory committee structure}

\cp

\p The \advcom must be formed within one month of the start of the student's
program.

\p The \index{supervisor} will select a provisional committee, in consultation
with the student. A \formGSP listing the \advcom members must be submitted to
the \GS.

\p Membership on an \advcom is flexible and may change during the course of a
student's program. Students must complete a \formGSPU when a decision to make
changes to committee membership is made. The \advcom structure differs for MSc
and PhD, as explained in sections~\ref{sec:msc_program_requirements} and
\ref{sec:phd_program_requirements}.

% above was: sections B and C

\p Advisory committees are formally sub-committees of the Department appointed
both to provide expert advice to students and to evaluate and report on student
progress.

\p All committee members must have Faculty of Graduate Studies membership (most
commonly as an \di{External Scholar}). Memberships should be applied for
through the \GS.

\p Committees are changed by mutual consent within the committee or by the
Chair of the committee and student in consultation with the \GC or Chair of the
Department. Limits to changes are set by the availability of particular people
as supervisors and of research facilities and funds (noting that most research
grants are awarded for specific projects).

%A9.1	ADVISORY COMMITTEE MEETINGS
%\subsection{Advisory committee meetings}
%
%\cp

\subsection{Advisory committee meeting schedule}
\cp
\p There must be no less than two advisory committee meetings held per academic
year\footnote{FGS Regulations, section 9.3: Supervisory Committees
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10852&loaduseredits=False}{\ExternalLink}.}

\p Advisory committee meetings may be held as often as required by the student
and/or supervisor. It is the responsibility of the student to ensure that
committee meetings are held. The supervisor will ensure that brief minutes of
these meetings are recorded and are filed with the \GS.

\p Following approval of the thesis proposal by the \advcom, the student should
have several meetings in addition to the regular committee meetings as follows:

\begin{itemize}

\item An interim progress meeting to verify that the research is on track.

\item A final progress meeting at which the committee agrees that sufficient
research has been conducted, and that thesis writing may proceed.

\item Individual meetings with advisory committee members to receive input on drafts or chapters.

\item A meeting at which the advisory committee agrees that the thesis is
defensible and that a defence may be scheduled. Alternatively, the advisory
committee may request further revision prior to approving a defence.

\end{itemize}


% %A9.2	COMMITTEE APPROVAL
% \subsection{Committee approval}
% 
% \cp

\p At several points in a student's program the approval of the advisory committee is essential:
\begin{itemize}
\item Initial discussion of research plan and direction; this should occur as early as possible in the student's program.
\item Approval of research plan and direction for thesis proposal (see sections~\ref{sec:msc_thesis_proposal}
    and~\ref{sec:phd_thesis_proposal}.
\item Acceptance of a thesis proposal.
\item Change of status, such as advancement to the PhD program (see section~\ref{sec:transfer_to_phd}).
\item Approval to write up sections of a thesis.
\item Approval of draft of thesis.
\item The acceptance of the thesis, in final form, as being ready for defence.
\end{itemize}


%B.	MSc  PROGRAM REQUIREMENTS

\section{\label{sec:msc_program_requirements}MSc program requirements}

% B1	MSc COURSE REQUIREMENTS

\subsection{\label{sec:msc_course_requirements}MSc course requirements}

\cp

\p MSc students must complete a minimum of 5 half-credits at the 5000-level or
higher, at least three of which must be chosen from the introductory core
courses (refer to section~\ref{sec:courses}).

\p Any student who anticipates a \index{transfer from MSc to PhD} transfer from
the MSc program to the PhD program should complete the course-work and
other requirements as listed in the PhD program requirements.

%B2	MSc COMMITTEE STRUCTURE
\subsection{MSc committee structure}

\cp
        
\p The Thesis \advcom must be formed within one month of the start of the
student's program.

\p The MSc \advcom consists of at least three members. There must be two full-time
faculty members from the Oceanography Department (not Adjunct Professors). If a
regular faculty member serving on the committee leaves the department, a
replacement with another faculty member must be made. One member will be from
another sub-discipline, and it is desirable that at least one member of the
committee be from outside the Department.


%B3	MSc THESIS PROPOSAL
\subsection{\label{sec:msc_thesis_proposal}MSc thesis proposal}

\cp

\p MSc students are expected to produce an approved \di[thesis
proposal!MSc]{thesis proposal} within one year of enrolling in the program.

\p The \di[thesis proposal!MSc]{thesis proposal} should be developed in
consultation with the supervisor and advisory committee. The scope of the
research should be such that it can be accomplished and the thesis written
within one year. Students should consult with their supervisors on the content
of a proposal.

\p The Supervisor will notify the Oceanography Office that the \di[thesis
proposal!MSc]{proposal} requirement has been satisfied and a copy of the
approved proposal must be placed on file with the \GC. MSc students may be
required to defend the research proposal at the discretion of their committee.

\p Because a large portion of a student's first year is consumed by course
work, the \di[thesis proposal!MSc]{thesis proposal} serves primarily as a
well-reasoned course of action, and not as an exhaustive literature review nor
as a deeply detailed description of methods. Fifteen pages of text should
provide ample space.


%B4	MSc THESIS
\subsection{MSc thesis}

\cp

\p Graduate students should pay particular attention to the Faculty of Graduate
Studies deadlines for submission of a thesis for graduation and fee payment
schedules.
 
\p The examination for the degree of Master of Science is subject to detailed
regulations of the Faculty of Graduate Studies\footnote{FGS Regulations,
section 10.3: Master's Theses \href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10853&loaduseredits=False}{\ExternalLink}.}.

\p The MSc thesis should report original research of such value as to merit
publication and be in a satisfactory and consistent literary form. Faculty of
Graduate Studies thesis formatting requirements are available on their
website\footnote{FGS Regulations, section 10.2: Preparation of Manuscript and
Submission of Theses
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=2&chapterid=399&topicgroupid=1435&loaduseredits=False}{\ExternalLink}.}.

\p The oral defence of the thesis is open to all members of the department and to
other interested persons. Notices of the thesis defence, including an abstract,
will be distributed to all members of the Oceanography Department and will be
posted.

%B4.1	MSc EXAMINING COMMITTEE
\subsection{MSc examining committee}

\cp

\p It is the supervisor's responsibility to select the examining committee
members, following the committee structure guidelines. The defence may
\emph{not} proceed in violation of the below composition of the examining
committee due to short-term absence of committee members.

\p The examining committee will consist of a minimum of 5 members:

\begin{itemize}
    \item 3 members of the advisory committee
        \begin{itemize}
            \item Supervisor(s)
            \item Full-time department faculty member
            \item One other member
        \end{itemize}

    \item Departmental Representative/Chair (from outside the
        committee)\footnote{The \GS will arrange for the Departmental
        Representative (normally the \GC or a designate). The Chair of the MSc
        thesis defence is independent of the examining committee. The
        Departmental Representative normally performs this duty.}

    \item External examiner from outside the advisory
        committee\footnote{Adjunct professors may fulfill the requirement of an
        external examiner.}

\end{itemize}


%B4.2	MSc THESIS DEFENCE TIMELINE
\subsection{MSc thesis defence timeline}

\cp

\p Students with immediate commitments subsequent to the defence must anticipate
the deadlines, rather than attempting to rush the defence process in a
compressed time frame.

\p Prior to proceeding with scheduling the defence, the thesis (in essentially
final form) should be distributed to the supervisor and committee to be
reviewed for suitability and final approval.

\begin{itemize}
\item Six weeks prior to defence:

        \begin{itemize}

            \item Have a format review done by FGS before the thesis is
                distributed to your examining committee

            \item Submit an MSc Examination Information Form to the Graduate
                Secretary

            \item Submit a Thesis Binding Submission Form and payment to the
                \GS

            \item Submit an electronic copy of the abstract to the Graduate
                Secretary

            \item Provide copies of the thesis to the supervisor for
                distribution to the examining committee

        \end{itemize}

    \item Four weeks prior to defence:

        \begin{itemize}

            \item Notify the \GS of any Audiovisual requirements

        \end{itemize}

    \item Two weeks prior to defence:
        \begin{itemize}

            \item Submit a printed copy of the thesis to the \GS

        \end{itemize}

    \item Following a successful Defence:

        \begin{itemize}

            \item Have Examining Committee Members sign the signature page

            \item Submit required changes to the supervisor within the
                specified timeframe

            \item Have a final format check of the thesis done by FGS

            \item Electronically submit the final version of the thesis via
                DalSpace\footnote{DalSpace
                \href{http://dalspace.library.dal.ca/}{\ExternalLink}.}

            \item Submit original completed forms to FGS (National Library of
                Canada Form, Title Page of Thesis, Signature page with original
                signatures, Copyright Page, Ethics Pages (if applicable),
                Student Contribution to Manuscripts (if applicable)

            \item Submit final copies of thesis to the \GS for binding (include
                copies of the signed signature page, and the Copyright Page
                with original signatures)

            \item Complete Exit Survey

        \end{itemize}

\end{itemize}

%B4.3	MSc THESIS DEFENCE OUTCOME
\subsection{MSc thesis defence outcome}

\cp

\p Outcomes: All theses are either approved or not approved. The categories are:
\begin{itemize}
\item Approve as submitted
\item Approved pending corrections and a clear timetable for completion (normally within one month)
\item Rejected but with permission to re-submit a revised thesis for re-examination with a clear timetable for completion (within one year)
\item Rejected outright
\end{itemize}

\p A simple majority determines the outcome, based on all the examiners except the
Departmental Representative, who will vote only in the event of a tie.

\p Following the defence, the candidate will receive a letter from the Chair of
the examining committee indicating the nature of any corrections to be made and
the time frame within which they are to be completed. In the event of an
unsuccessful defence, an explanation of the negative outcome is provided. A
copy of the report will be provided to the \GS.

\subsection{\label{sec:transfer_to_phd}Transfer to PhD}

\cp

\p A request to \index{transfer from MSc to PhD} transfer from the MSc program to the PhD program should be
completed no later than one month prior to the anniversary of the student's
admission date.

\p An MSc student may transfer to a PhD program without completing a MSc. Students
wishing to transfer should provide a letter to their advisory committee stating
the reason for the transfer, and justification that they are qualified for the
PhD program. A one-page explanation is sufficient. The student's \advcom is
required to meet, and the committee may recommend advancement to the PhD
program. A letter from the supervisor, justifying the advancement of the
student to the PhD program, should be submitted to the \GC along with the
student's letter no later than one month prior to the anniversary of the
student's admission date. The \gocom will then review the request.

\p The decision categories for the transfer to the PhD are:
 \begin{itemize}
     \item Approved
     \item Not approved
     \item Deferred
 \end{itemize}
%\p The decision categories for the transfer to the PhD are: ``Approved,'' ``Not
%approved,'' and ``Deferred.''
% \begin{itemize} \item Approved \item Not approved \item Deferred
% \end{itemize}

\p Students anticipating this change should take the required core courses in the
first year and take the qualifying examinations (see
section~\ref{sec:phd_qualifying_examination}) in the first year following
completion of the required core classes.

% above referred to C3 for QE

%C.	PhD  PROGRAM REQUIREMENTS
\section{\label{sec:phd_program_requirements}PhD program requirements}

%C1	PhD COURSE REQUIREMENTS
\subsection{\label{sec:phd_course_requirements}PhD course requirements}

\cp

\p PhD students must complete a minimum of 6 half-credits at the 5000-level or
higher, at least two of which must be chosen from the introductory core courses
outside the student's sub-discipline (refer to section~\ref{sec:courses}).

% above referred to A2.1 originally

\p Courses are to be chosen in consultation with the advisory committee. The
various sub-disciplines have individual course requirements, some of which
exceed the minimum requirements, as explained below.

%C1.1	BIOLOGICAL OCEANOGRAPHY GUIDELINES
%\subsubsection{Biological Oceanography guidelines}

\p The normal course requirements for a PhD in Biological Oceanography are:
\begin{itemize}
\item Core Classes: Biological Oceanography (5140), Chemical Oceanography (5130), and Physical Oceanography (5120)
\item A minimum of one other half credit drawn from Geological Oceanography (5110) or a suitable Atmospheric Science course
\item Other Classes, as required by the Advisory Committee
\end{itemize}

%C1.2	CHEMICAL OCEANOGRAPHY GUIDELINES
%\subsubsection{Chemical Oceanography guidelines}

\p The normal course requirements for a PhD in Chemical Oceanography are:

\begin{itemize}
\item Core Classes: Chemical Oceanography (5130), Geological Oceanography (5110), Physical Oceanography (5120), and Biological Oceanography (5140)
\item Advanced Chemical Oceanography (modular) (5290)\fixme{5290 not offered since 2007/9}
\item Other Classes, as required by the Advisory Committee
\end{itemize}

%C1.3	GEOLOGICAL OCEANOGRAPHY GUIDELINES
%\subsubsection{Geological Oceanography guidelines}

\p The normal course requirements for a PhD in Geological Oceanography are:

\begin{itemize}
\item Core Classes: Geological Oceanography (5110) and two other Introductory Classes.
\item Other Classes, as required by the Advisory Committee
\end{itemize}

%C1.4	PHYSICAL OCEANOGRAPHY GUIDELINES
%\subsubsection{Physical Oceanography guidelines}

\p The normal course requirements for a PhD in Physical Oceanography are:

\begin{itemize}
\item Core Classes: Physical Oceanography (5120) and two other Introductory Classes.
\item Advanced Classes: Fluid Dynamics (5311), Time Series Analysis (5210),
Ocean Dynamics (5221), Estuary, Coast and Shelf Dynamics (5222).
\item Other Classes, as required by the Advisory Committee e.g. Numerical
Modelling (5220), Ocean Waves (5223), Introduction to Acoustical Oceanography
(5250), Advanced Marine Particles (5293).
\end{itemize}

\fixme{I don't think we really should say 'advanced' vs 'other'. The waves
class is surely advanced. And time series is much more ``other'' than
advanced, since it's not tied to oceanography.}

%C2	PhD COMMITTEE STRUCTURE
%\subsection{PhD committee structure}

\p The Thesis Advisory Committee must be formed by October 1st of the first year
of study.

\p The PhD Advisory Committee consists of at least four members. There must be two
full-time faculty members from the Oceanography Department (not Adjunct
Professors) on the advisory committee. If a regular faculty member serving on
the committee leaves the department, a replacement with another faculty member
must be made.

\p One member will be from another sub-discipline, and it is desirable that at
least one member of the committee be from outside the Department.


%C3	PhD QUALIFYING EXAMINATION
\subsection{\label{sec:phd_qualifying_examination}PhD Qualifying Examination}

\cp

\p The \di{qualifying examination} should be completed between months 9 and 12 of
the program.  Students transferring from the MSc program should take the
qualifying examinations within 12 months of their transfer.

\p This examination is designed to assess the student's background knowledge, with
the primary aim of identifying weaknesses that need to be addressed in order
for a student to undertake research at the Ph.D. level in the chosen field. The
format varies with sub-discipline. In each case there is an oral component, but
in some cases there can also be a written assignment before or after the oral
examination.

\p The \curcom  may grant extensions up to month 15 if the student and supervisor
document a compelling conflict, e.g. \di{fieldwork}.
 

%C3.1	QUALIFYING EXAMINATION COMMITTEE
\subsubsection{Qualifying Examination committee}

\cp

\p The examination committee has at least 4 members, including a faculty member
from outside the sub- discipline, and a departmental representative who acts as
chair. The \GS will arrange for the Departmental Representative (normally the
\GC or a designate). The composition of the rest of the examining committee is
set by the sub-discipline.

\p The topics of the examination are student-specific, and may include items of a
general oceanographic nature (tailored to the courses taken) as well as items
that are more tightly focused on the sub-discipline and the intended area of
research.

%C3.2	FORMAT AND GUIDELINES
\subsubsection{Format and guidelines}

%C3.2.1  BIOLOGICAL OCEANOGRAPHY QE GUIDELINES

\input{BO_Qualifying_Aug2013_draft.tex}

\paragraph{Biological Oceanography QE guidelines}\hfill

\cp
\p The core of the qualification process is an oral examination based on a reading
list tailored to the student's research area. In some cases this may be
followed by a written examination. The details and the timing are as follows.

\begin{enumerate}

\item The PhD candidate prepares a brief (2-3 page) summary of his or her
research interests and sends it to the BO faculty. The student's summary of
research interests guides faculty members in the selection of papers; it will
not be used as source material during the exam, and neither the student nor the
committee members bring copies of the research summary to the exam.

\item Based on this document, at least 3 of the available BO faculty members
(including the supervisor) each contribute one or two papers or book chapters
to a reading list. These papers are intended to serve as guides for the exam
and to provide entry points for discussions of oceanographic topics during
questioning. The process is organized by the research supervisor, who vets the
papers to ensure appropriateness, and then communicates the list to the student
and the BO faculty within 2 weeks of receipt of the student's research summary.

\item It is the responsibility of the candidate's supervisor to: a) form the
Examining Committee; and b) schedule the exam. The committee will include a
Chair, at least three members of the BO faculty, and one Oceanography faculty
member from out of the BO sub-discipline, preferably in the sub-discipline most
closely related to the student's research. The Chair (a member of the
Oceanography faculty not in BO) will serve as moderator of the exam and will
not ask questions. 3 members of the BO group must approve the form of the
examining committee no later than 3 weeks prior to the date of the exam. The
composition of the committee will be revealed to the student and the student
will see that each member has copies of the readings.

\item Within 6 weeks after the reading list is given to the student, an oral
examination is held.

\item The exam begins with a 20-minute presentation by the student,
highlighting aspects of the papers or chapters that are relevant to his or her
research interests.

\item During the next 1-1.5~h, the committee asks questions about oceanography
and topics related to the students research interests. Entry points into
questions and discussions are to be based on the reading material that was
provided to the student and by the student's presentation.

\item After the exam, the committee meets in camera to agree on comments and
recommendations to the student and to the students committee, as appropriate.

\item The possible outcomes of the oral examination are:
\begin{itemize}
\item The candidate passes without extra conditions.

\item The candidate passes, but is informed of weaknesses that should be
addressed during the PhD work, e.g. in courses or in directed studies.

\item The candidate is required to take a written examination, at a date
determined during the oral examination meeting. This examination will be based
on the topics that arose during the oral examination and will not exceed three
hours. The committee can then pass the student with no extra conditions, or be
informed of weaknesses that should be addressed during the PhD work, e.g. in
courses or in directed studies.

\end{itemize}

\item The Chair is to report in writing to the Chair of the Curriculum
Committee on the outcome of the exam, with copies to candidate and examining
committee.

\end{enumerate}


%C3.2.2  CHEMICAL OCEANOGRAPHY QE GUIDELINES
\paragraph{Chemical Oceanography QE guidelines}\hfill

\cp

\p No discipline-specific guidelines--follow general Departmental guidelines.

%C3.2.3  GEOLOGICAL OCEANOGRAPHY QE GUIDELINES
\paragraph{Geological Oceanography QE guidelines}\hfill

\cp

\p No discipline-specific guidelines--follow general Departmental guidelines.

%C3.2.4  PHYSICAL OCEANOGRAPHY QE GUIDELINES
\paragraph{Physical Oceanography QE guidelines}\hfill
\cp

\p The core of the qualification process is an oral examination based on a reading
list tailored to the student's research area. In some cases this may be
followed by a written examination. The details and the timing are as follows.

\p The PhD candidate prepares a brief (1 page) description of the intended area of
research and sends it to the physical oceanography faculty members 5 weeks
prior to the oral examination.

\p Based on this document, each available faculty member in the sub-discipline
contributes a paper to a reading list. This process is organized by the
research supervisor, who then communicates the list to the student and the
Physical Oceanography faculty members within 1 week of receipt of the student's
research summary.

\p Within 4 to 6 weeks, an oral examination is held, to test the student's
comprehension of the reading list and its relevance to the intended research.
The examination committee comprises all available faculty in the
sub-discipline, a faculty member from another sub-discipline, and a
Departmental representative.

\p The possible outcomes of the oral examination are:
\begin{itemize}
\item The candidate passes without extra conditions

\item The candidate passes, but is informed of weaknesses that should be
addressed during the PhD work, e.g. to be addressed in courses or in directed
studies

\item The candidate is required to sit written examination, at a date
determined during the oral examination meeting. This examination will be based
on the topics arising during the oral defence.

\end{itemize}



%C3.3	QUALIFYING EXAMINATION OUTCOME
\subsubsection{Qualifying Examination outcome}

\cp
\p There are several possible outcomes of the qualifying examination:

\begin{itemize}

\item Continuation in the PhD program with additional course work or directed studies
\item Continuation in the PhD program without additional course work or directed studies
\item Transfer to the MSc program
\item Transfer to the PhD program with additional course work or directed studies
\item Transfer to the PhD program without additional course work or directed studies
\item Continuation in the MSc program

\end{itemize}

% C4	PhD THESIS PROPOSAL AND DEFENCE
\subsection{\label{sec:phd_thesis_proposal}PhD thesis proposal and defence}

\cp

\p PhD students are expected to complete the \di[thesis proposal!PhD]{thesis
proposal} and examination of the proposal within 20 months of enrolling in the
graduate program.

\p The advisory committee sets the exact timing.

\p In order to proceed with the \di[thesis proposal!PhD]{proposal}, the student
will have completed the required first year courses, have a named supervisor
and advisory committee, and received approval from the supervisor.

\p The requirement consists of a written \di[thesis proposal!PhD]{thesis
proposal} and a public oral defence.  The written proposal will include an
overview of the research topic and relevant background material, and a plan for
carrying out the PhD research.  The oral portion is a defence of the proposal,
including other relevant areas of Oceanography. The topics covered in the
defence may relate not just to the planned research, but also to wider
oceanographic issues, as deemed appropriate by the defence committee.

\p The defence is open to attendance by all interested persons, except for the
in-camera portion in which the decision is made by the examining committee.
Notices of the thesis proposal defence, including an abstract will be
distributed to all members of the Oceanography Department and will be posted.


%C4.1	EXAMINING COMMITTEE
\subsubsection{Examining committee}

\cp

\p The examining committee is selected by the supervisory committee, and
consists of members from that committee, external members, and a departmental
representative. The \GS will arrange for the Departmental Representative
(normally the \GC or a designate).

\p The Chair of the PhD proposal defence is independent of the examining
committee. The Departmental Representative normally performs this duty.

\p Students must make the proposal available in the Department of Oceanography
office at least 10 working days prior to the oral defence.  The proposal
typically receives input from the research supervisor or others prior to its
submission. Approval from the advisory committee is required to proceed with
the oral defence.

%C4.2	PROPOSAL TIMELINE
\subsubsection{Proposal timeline}

\cp

\p It is the student's responsibility to initiate the proposal process and develop
the proposal in coordination with the supervisor and committee. Two terms are
more than adequate to complete this requirement. The scheduling of the proposal
defence should be planned well in advance so that it can be carried out prior
to the 20th month of the program.

\p There is no restriction on early completion of the proposal and oral exam (e.g.
in the summer between first and second years or fall term of second year).

\p Because the proposal is a formal requirement of the Department, it carries the
same weight as courses, thesis defence, etc. Failure to complete the proposal
within the required time frame reflects poorly on the student and can endanger
standing in the program.


%C4.3	PROPOSAL OUTCOME
\subsubsection{Proposal outcome}

\cp

\p Based on both the written proposal and oral defence, the supervisor will
notify the \GS, \GC, student, and examining committee in writing of one of the
four possible outcomes:

\begin{itemize}
\item Continuation in the PhD program
\item Permission to re-defend
\item Transfer to the MSc program
\item Withdrawal from the program
\end{itemize}

\p Revisions, and/or a re-defence must be completed prior to the end of the exam
period in the winter term. Following the defence, a copy of the approved
proposal must be placed on file in the Oceanography office.


%C5	PhD THESIS
\subsection{PhD thesis}

\cp

\p The PhD thesis should report original research of high caliber carried out by
the student. It should be of such value as to merit publication and be in
satisfactory literary form, with clearly presented figures and other supporting
material. The thesis should be presented first, in draft form, to the
supervisor who will provide initial comments and approve distribution to the
rest of the committee. Typically, the writing phase is very intensive, and
involves working closely with the supervisor and committee members. In many
cases, the student writes journal articles that link closely with the thesis
chapters, and the Faculty of Graduate Studies has conventions on how this is to
be handled. As the thesis reaches completion, the advisory committee will meet
to discuss any final changes to the thesis and sign the PhD Thesis Submission
Form; this is just one of a series of steps that the student must keep in mind
(see section~\ref{sec:phd_thesis_defence_timeline}). 


%C5.1	PhD EXAMINING COMMITTEE
\subsubsection{PhD examining committee}
 
\cp

\p It is the supervisor's responsibility to select the examining committee
members, following the committee structure guidelines.

\p The examining committee will consist of a minimum of 5 members:

\begin{itemize}

\item 3 members of the advisory committee

    \begin{itemize}

        \item Supervisor(s)

        \item Full-time department faculty member

        \item One other member

    \end{itemize}

\item A Departmental Representative, from outside the committee.  The Graduate
    Secretary will arrange for the Departmental Representative, normally the
    Graduate Coordinator or a designate.

\item External examiner (from outside the University)

    \item Chair\footnote{The Chair of the PhD thesis defence is independent of
        the examining committee. The Faculty of Graduate Studies will appoint
        the Chair.}

\end{itemize}


%C5.2	PhD THESIS DEFENCE TIMELINE
\subsubsection{\label{sec:phd_thesis_defence_timeline}PhD thesis defence timeline}

\cp

\p Students should be aware of the time scales required by FGS and Departmental
regulations, which stretch over a period of more than eight months before the
defence.

\p Prior to proceeding with scheduling the defence, the thesis (in essentially
final form) should be distributed to the supervisor and committee to be
reviewed for suitability and final approval.


\begin{itemize}
    \item Twelve weeks prior to defence:

        \begin{itemize}

            \item Submit Request to Arrange an Oral Defence Form, listing 3
                choices for external examiners along with CV for external
                examiner of first choice to the \GS.

        \end{itemize}

    \item Six weeks prior to defence:

        \begin{itemize}

            \item Have a format review of the thesis done by FGS before it is
                distributed to the examining committee

            \item Submit a PhD Thesis Submission From and copy of the thesis
                for the External Examiner to the \GS 

            \item Submit a PhD Examination Information Form to the Graduate
                Secretary

            \item Submit a copy of the student's CV to the \GS 

            \item Submit a Thesis Binding Submission Form and payment to the
                \GS

            \item Submit an electronic copy of the abstract to the Graduate
                Secretary

            \item Provide copies of the thesis to the supervisor for
                distribution to the internal examining committee members

        \end{itemize}

    \item Four weeks prior to defence:

        \begin{itemize}

            \item Notify the \GS of any Audiovisual requirements

        \end{itemize}

    \item Two weeks prior to defence:
        
        \begin{itemize}

            \item Submit a printed copy of the thesis to the \GS 

        \end{itemize}


    \item Following Defence:

        \begin{itemize}

            \item Have Examining Committee Members sign the signature page

            \item Submit required changes to the thesis within the specified
                timeframe

            \item Have a final format check done by FGS

            \item Electronically submit the final version of the thesis via
                DalSpace\footnote{DalSpace
                \href{http://dalspace.library.dal.ca/}{\ExternalLink}.}.

            \item Submit original completed forms to \FGS (National Library of
                Canada Form, Title Page of Thesis, Signature page with original
                signatures, Copyright Page, Ethics Pages (if applicable),
                Student Contribution to Manuscripts (if applicable)

            \item Submit final copies of thesis to the \GS for binding (include
                copies of the signed signature page, and the Copyright Page
                with original signatures)

            \item Complete Exit Survey

        \end{itemize}

\end{itemize}

\p Some of the above deadlines are set by the Oceanography Office. See the FGS
website for advice on preparing for a Doctoral defence\footnote{FGS Theses and
Defences
\href{http://www.dal.ca/faculty/gradstudies/currentstudents/thesesanddefences/defense.html}{\ExternalLink}.}.


%C5.3	PhD THESIS DEFENCE OUTCOME
\subsubsection{PhD thesis defence outcome}

\cp

\p The student opens the examination with a 20-minute presentation of the thesis.
Other faculty, students and public may be present, and ask questions, but the
critical decisions are made in private and are the responsibility of the
examining committee.

\p Following the defence, the candidate will receive a letter from the Faculty of
Graduate Studies indicating the nature of any corrections to be made and the
time frame within which they are to be completed.  In the event of an
unsuccessful defence, an explanation of the negative outcome is provided.  

%C5.4 PhD SEMINAR PRESENTATION
\subsubsection{PhD seminar presentation}

\cp

\p A departmental seminar on the thesis should be presented after successful
completion of the thesis examination.

%D.	FINANCIAL  SUPPORT
\section{Financial  support}

\cp

\p Dalhousie Graduate Fellowships will normally be provided for two years for an
MSc student and four years for a PhD student.

\p Scholarship support is provided directly to some students by the Natural
Sciences and Engineering Research Council (\di{NSERC}), the \di{Killam
Foundation} through Dalhousie, by companies, and by other agencies, both
national and international. Students are encouraged to apply for external
support whenever possible.

\p Full-time students without such special scholarship support are generally
awarded Dalhousie Graduate Fellowships from funds that come partly to the
Department from the University and mainly from individual faculty members'
research grants.

\p Students may also receive extra support by taking on a teaching assistantship,
or by demonstrating in undergraduate laboratories.

\p Students should discuss any additional jobs with their advisory committee to
ensure that the time commitments outside the Department are not excessive, and
do not interfere with program expectations.


%D1	FGS CONFERENCE TRAVEL GRANT
%\subsection{FGS conference travel grant}


\p
Graduate students are eligible to apply for one \di{travel grant} per degree at
Dalhousie (Master's or Doctoral) through the \FGS\footnote{FGS Conference
Travel Grant Application Form and Guidelines
\href{http://www.dal.ca/faculty/gradstudies/funding/grants.html}{\ExternalLink}.}.
Applications must be submitted through the \GS at least one month prior to the
date of travel.  Students must be registered in a graduate program at the time
of application and at the time of the conference. In order to be eligible,
students must present a poster of paper based on the results of their graduate
thesis research at a national or international scholarly meeting or conference.
Departmental approval must be given to these applications.


%E.	INTELLECTUAL  PROPERTY
\section{Intellectual  property}

\cp

\p The Faculty of Graduate Studies is developing\fixme{probably already exists
-- check into it}  a policy on Intellectual Property, which will be available
on the Faculty of Graduate Studies website.  If students and/or faculty have
concerns or doubts about any issue pertaining to any part of Section 6, consult
with your Chair, \GC, or Supervisor, or contact the Faculty of
Graduate Studies for advice. If you feel uncomfortable with approaching your
immediate supervisor, then go to the next level and ask to be heard in
confidence\footnote{FGS Regulations, section 6.4: Policy on Intellectual
Property
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10849&loaduseredits=False}{\ExternalLink}.}.

%\elink{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10849&loaduseredits=False}, section 6.4.}.

%E1	ACADEMIC INTEGRITY
\section{Academic integrity}

\cp

\p Any material submitted by a student at Dalhousie University may be checked
for originality to confirm that the student has not plagiarized from other
sources.  \di[plagiarism]{Plagiarism} is a serious academic offence that may
lead to a failing grade, suspension or expulsion from the University, or even
the revocation of a degree.  It is essential that there be correct attribution
of authorities from which facts and opinions have been derived. Prior to
submitting any paper in a course, or material for a thesis, students should
read the Policy on Integrity in Scholarly Activity\footnote{FGS Regulations,
section 6.3: Policy on Integrity in Scholarly Activity
\href{http://academiccalendar.dal.ca/Catalog/ViewCatalog.aspx?pageid=viewcatalog&catalogid=58&chapterid=2678&topicgroupid=10849&loaduseredits=False}{\ExternalLink}.}.

%\include{appeals}


\section{Procedures for Student Academic Appeals}
%and Governance of the Ad Hoc Appeal Committee}

\subsection{Introduction}

\cp

\p The procedures detailed below will be followed in cases where a student
wishes to appeal an academic procedure, other than a grade, on the set of
criteria detailed in section~\ref{sec:basis_for_an_appeal}. (Grade appeals are
handled by the University's Registrar.)

\p The regulations and procedures given below cover appeals launched by
both undergraduate and graduate students in the Department of Oceanography at
Dalhousie University.  

\p These procedures do not apply to labor disputes between an employee
(be that person a student, post-doc, etc.) and an employer (faculty, staff) in
regard to working conditions.  Persons wishing to dispute a labor condition
should consult either the Chair of the Department and/or their union agreement,
if applicable, or the laws governing labour relations in Nova Scotia and the
procedures detailed therein. 

\p These regulations do not apply if the cause of the dispute centers
on discrimination based on gender, race, nationality, sexual orientation,
gender identity, official language, or disability.  The complainant in such
cases must contact the Dalhousie Office of Human Rights, Equity \& Harassment
Prevention for procedures in such cases.  Nor are these regulations applicable
to cases of sexual harassment, assault, criminality, etc., where Dalhousie
Security or Law Enforcement should be involved.  

\p There are no appeals on admission decisions, including transfers to
the PhD program.  

%\end{document}

\subsection{\label{sec:basis_for_an_appeal} Basis for an Appeal}

\cp

\p The procedures below apply to the administration of qualifying and
preliminary examinations, comprehensive examinations, thesis proposal defenses,
and Master's thesis defenses.  Results (grades) of such examinations cannot be
appealed as oral portions cannot be regraded.  

\p Appeals related to Doctoral defenses must be directed to Faculty of
Graduate Studies in the first instance. 

\p Appeals of course examination grades (re-grading) are direct through
the University Registrar's Office.
%, and not the Department of Oceanography.
%Some clarification on the method of reassessment is given below in
%section~\ref{sec:grade_reassessment}; otherwise the appellant must follow the
%procedures set by the Registrar. \fixme{should we just leave the rules to the
%registrar, thus axiomatically avoiding contradiction?}
 
\p The grounds for appeal are limited to the following.  For all these grounds,
the onus is on the appellant to prove, through factual documentation, that such
conditions existed. 

%\end{document}

\begin{enumerate}

    \item Procedural Unfairness.  (This means that the method used to
        administer the academic process, e.g., an exam, was unreasonably and
        inherently stacked against any person attempting the process.  Note:
        exams and their content, in and of themselves, are not procedurally
        unfair.  Differences in how an exam is administered between the
        sub-disciplines in the Department also does not constitute unfairness,
        as long as the method is applied consistently within that
        sub-discipline.)

    \item Bias. (This means that the process was conducted in such a way as to
        disadvantage the particular appellant, relative to other persons
        undertaking the same process.)

    \item Irregularity in Procedure. (This means that the process did not
        follow the procedures set in the regulations and guidelines governing
        the actions of the Department and its faculty.)

    \item Inappropriate or Unfair Expectation(s). This means that the
        deliverable of the process could not reasonably be expected of a person
        who undertakes assiduously, genuinely, and accurately the steps leading
        to completion of that process.) 

\end{enumerate}

% ``Procedural Unfairness'' means that the method used to administer the academic
% process, e.g., an exam, was unreasonably and inherently stacked against any
% person attempting the process.  Note: exams and their content, in and of
% themselves, are not procedurally unfair.  Differences in how an exam is
% administered between the sub-disciplines in the Department also does not
% constitute unfairness, as long as the method is applied consistently within
% that sub-discipline.  
%
% ``Bias'' means that the process was conducted in such a way as to disadvantage the
% particular appellant, relative to other persons undertaking the same process.    
%
%``Irregularity in procedure'' means that the process did not follow the
%procedures set in the regulations and guidelines governing the actions of the
%Department and its faculty.  
%
%``Inappropriate or unfair expectation'' means that the deliverable of the
%process could not reasonably be expected of a person who undertakes
%assiduously, genuinely, and accurately the steps leading to completion of that
%process.   


\p A written appeal must be submitted to the Chair of the Department of
Oceanography within 30 days following the event or circumstances being
appealed. 

\p The appeal submission must include:
\begin{enumerate}

\item A description of the exact nature of the appeal, including a summary of
    events and chronology

\item Specific details of the alleged unfairness, bias or irregularity, and any
    other relevant consideration or information

\item The requested resolution of the appeal, which is limited to a reasonable
    academic action(s). 

\end{enumerate}

\p The submission of an appeal will engender the following actions by
the Chair of the Department:

\begin{enumerate}

\item The Chair will contact both the appellant and the person responsible for
the process being appealed to see if an informal resolution is first possible, and

\item Failing an informal resolution, the Chair will constitute an Ad Hoc
Appeal Panel for that specific appeal.  The nature of and procedures for that
Panel are detailed below in section~\ref{sec:the_ad_hoc_appeal_panel}.

\end{enumerate}

\p Decisions of the Ad Hoc Appeal Panel are forwarded to the Chair, who
will communicate the decision in writing to the appellant.

\p Decisions of the Ad Hoc Appeal Panel are subject to further appeal to
the Faculty of Science (undergraduates) or the Faculty of Graduate Studies
(graduate students), or Senate, as regulations in those administrative units
specify. 

%\end{document}


\subsection{\label{sec:the_ad_hoc_appeal_panel}The Ad Hoc Appeal Panel}

\cp

\p Decisions on appeals are made by an Ad Hoc Appeal Panel.

\p The Panel does not have fixed membership.  Instead, it will consist
of three faculty members, chosen on a fixed rotating basis from all the faculty
in the Department.  The professor(s) responsible for the process being appealed
will be excluded.  For graduate students, the student's supervisor or
co-supervisors, or for undergraduates, the student's honors supervisor (if
applicable) will also be excluded from the list.

\p An appropriate graduate student (for undergraduate appeals) or
Post-doctoral fellow (for graduate appeals) will constitute the fourth and
final member of the Panel.  Any student on the Panel will not have a personal
or academic conflict of interest with respect to the appellant.  For
undergraduates, the student will be chosen from the graduate students.  For
graduate students, the student member will be chosen from the Post-doctoral
Fellows, without conflict of interest.  

\p The Panel will examine the appeal and determine if a \emph{prima
facie} case exists for the appeal.  (\emph{Prima facie} in this context means
the appeal document contains sufficient evidence to support the stated claim.)
If such a case exists, then the Panel will convene a formal Appeal Hearing at
the first opportunity when all involved, i.e., appellant, the faculty member
who owns the process, and the Panel members can meet to hear verbal arguments
on the appeal, but preferably within three weeks of constitution of the Panel.
If there is no \emph{prima facie} case, the Panel will inform the Chair and the
appeal will be dismissed at the Departmental level.

\p If a Hearing is convened, the Panel will chose a Chair who will
control and direct all communication.

\p The Chair of the Panel will make the faculty owner of the process
aware of the appeal and provide, in a timely manner, a copy of the appeal
document to the process-owner (faculty member).  

\p Before the Hearing, the process owner will provide the Panel with a
written rebuttal to the appeal within 5 working days of the appeal being
received.  (Exceptions will be made to accommodate persons being at sea or
subject to other serious limitations in availability.)  The Panel will forward
that rebuttal to the appellant.


\subsection{The Appeal Hearing}
\cp

\p The Hearing will adhere to the following steps:

\begin{enumerate}

\item The Chair of the Panel will direct the flow of the Hearing, making sure
    both parties are properly heard, and directing questioning;

\item The Chair of the Panel will first invite the appellant to present and
    explain their case; this will be followed by an invitation to the faculty
    member owner for a rebuttal;

\item Neither the appellant nor the owner can ask questions during these
    presentations; however, any Panel member may interrupt the presentation to
    ask questions;

\item At the end of the presentations, the appellant will be invited to direct
    questions to the faculty owner for the purpose of seeking clarifications
    (only); after that is completed, the faculty owner will be invited to do
    the same;  the Panel members may interrupt these questioning sessions at
    any time to ask their own questions; the Chair of the panel will ensure
    that this questioning remains civil and on topic;

\item At the end of the questioning, the Chair of the Panel will ask the
    appellant and faculty owner to leave and convene an \emph{in-camera}
    discussion of the presentations and Appeal Document, the Panel will decide
    on the validity of the appeal and of requested resolution.

\end{enumerate}

\p The Chair of the Panel will communicate, in writing, the decision of the
Panel within 3 working days (except for serious time conflicts such as noted
those mentioned at \ref{sec:the_ad_hoc_appeal_panel}.7 above).\fixme{cross ref}

\p In an appeal process, the student has the right to representation.
The student is required to inform the Chair of the Panel, in writing, if s/he
will have a representative at the appeal, or intends to call witnesses. 

\p Witnesses may be called by either party, but only if those witnesses
can testify about direct knowledge of the matter.  No character or indirect
testimony is permitted.  The Panel Chair must be informed 72-hours before a
Hearing if a witness(es) will be called, along with that person's identity and
significance to the Hearing.  The Chair of the Panel will then inform the other
party of the witness(es).  

\p No persons providing solely moral support can be present at a
Hearing.  A person providing aid to someone with a disability is permitted, as
is a translator, but such persons cannot testify at a Hearing.

%\end{document}

\subsection{Supplementary Points}

\cp

\p The Ad Hoc Appeal Panel has no jurisdiction to hear student appeals
on a matter involving a requested exemption from the application of
Departmental, Faculty or University regulations or procedures, except when
irregularities or unfairness in the application thereof is alleged.  This means
that only procedural issues, and not the merits of the regulations, are subject
to appeal.

\p The Ad Hoc Appeal Panel may not render decisions counter to
Departmental, Faculty or University regulations, nor can it make decisions that
go beyond strictly academic matters, e.g., financial or administrative.  If the
requested resolution contains such points, these must be ignored and dismissed.   

\p Matters involving allegations of ``failure to supervise'' by a
graduate student will be referred directly to the Faculty of Graduate Studies
for resolution.  


\addcontentsline{toc}{section}{Index}
\printindex

\end{document}
